### 2014-05-21 intro.tex
Finish intro?

By ``showing our work'' clearly enough for a computer to create the
design, tweaking a protocol's assumptions becomes a matter of changing
the inputs to a computer program and running it again. My colleagues
and I have found that this approach yields better peformance than
conventional protocols, while allowing adjacent layers to evolve more
freely. It's not clear when computer-generated protocols will be
practically deployed on the broad Internet, but what computers can
teach us about protocol design can be applied to help humans create
better protocols now.

@@

On a simulated 15~Mbps link with eight senders contending and an RTT
of 150~ms, a computer-generated congestion-control algorithm achieved
the following improvements in median throughput and reductions in
median queueing delay over these existing protocols:

@@

\subsection{Algorithms should collect rate estimates as signals}

@@

\vspace{\baselineskip}
\enlargethispage{\baselineskip}

@@

\textcolor{white}{.}\hspace{\parindent} A Remy-generated algorithm, or
RemyCC, measures the \textbf{rate} of acknowledgment arrivals, and the
\textbf{rate} of transmission implied by the echoed sender timestamps
in those acknowledgments. We can quantify the value of these signals
by removing one of them and rerunning the optimization process. The
results are markedly worse with either signal removed.  From
Figure~\ref{f:tpdelaydb4} (a 15~Mbps link with 8 contending nodes):
%}

@@

More broadly, automated protocol-design tools have led us to a style
of design where the value of information can be evaluated
empirically. We may give Remy access to a new congestion
signal---either a function of existing TCP acknowledgments from the
receiver, or based on a new source of information---and measure
how much the resulting congestion-control algorithm improves over the
status quo. Similarly, we can knock out an existing signal and rerun
Remy to measure how much the loss would hurt performance.

After investigating many candidate signals, we have ended up with
computer-generated algorithms that collect four congestion signals
(Section~\ref{ss:signals}). \textbf{None of these signals} is
collected by existing TCP congestion-control schemes. Our
simulation experiments indicate each signal is individually valuable on
top of the other three.

@@

@@

>>>>

By ``showing our work'' clearly enough for a computer to create the
design, tweaking a protocol's assumptions becomes a matter of changing
the inputs to a computer program and running it again. My colleagues
and I have found that this approach yields better performance than
conventional protocols, while allowing adjacent layers to evolve more
freely. It's not clear when computer-generated protocols will be
practically deployed on the broad Internet, but what computers can
teach us about protocol design can be applied to help humans create
better protocols now.

@@

On a simulated 15~Mbps link with eight senders contending and a
round-trip time (RTT) of 150~ms, a computer-generated
congestion-control algorithm achieved the following improvements in
median throughput and reductions in median queueing delay over these
existing protocols:

@@

\subsection{Algorithms could benefit from collecting rate estimates}

@@

\vspace{0.85 \baselineskip}
\enlargethispage{\baselineskip}

@@

\textcolor{white}{.}\hspace{\parindent} A Remy-generated algorithm, or
RemyCC, measures the \textbf{rate} of acknowledgment arrivals, and the
\textbf{rate} of transmission implied by the echoed sender timestamps
in those acknowledgments. We can quantify the value of these signals
by removing one of them and rerunning the optimization process. The
results are markedly worse with either signal removed.  From
Figure~\ref{f:tpdelaydb4}:
%}

@@

\newpage

More broadly, computer-generated protocols have led us to a style
of design where the value of information can be evaluated
empirically. We may give Remy access to a new congestion
signal---either a function of existing TCP acknowledgments from the
receiver, or based on a new source of information---and measure
how much the resulting congestion-control algorithm improves over the
status quo. Similarly, we can knock out an existing signal and rerun
Remy to measure how much the loss would hurt performance.

After investigating many candidate signals, we have ended up with
computer-generated algorithms that collect four congestion signals
(Section~\ref{ss:signals}). \textbf{None of these signals} is
collected by traditional TCP congestion-control schemes. Our
simulation experiments indicate each signal is individually valuable on
top of the other three.

@@

Traditional congestion-control algorithms are window-based and
ack-clocked.~\cite{Srikant} Some proposed algorithms use
\emph{pacing}---transmitting packets according to an internal clock at
the sender---but the performance of these algorithms can display
counter-intuitive weaknesses.~\cite{understandingpacing}

After studying the behavior of Remy's algorithms and examining their
internal logic, we found they often behave as a fine-grained
hybrid of these two techniques, qualitatively different in behavior
than prior window- or pacing-based algorithms. In steady state, the
RemyCCs alternate between windowing and
pacing behavior. Within any interval that is at least 2 RTTs, some of
the algorithm's transmissions will have been gated by a windowing
constraint (waiting until the congestion window is larger than the
number of packets in flight), and some will have been gated by a
pacing constraint (waiting until the last packet transmitted was sent
long enough ago).

In practice, this hybrid behavior tends to produce quicker convergence to a
fair allocation of network resources when a flow starts or finishes
than traditional TCP-like schemes. We suspect there may be more to say
on the topic of ``windowing vs.~pacing'' if new algorithms similarly
attempt to use both techniques nearly concurrently.

\subsection{TCP-friendliness carries a benefit, but also a cost}

Congestion-control protocols are often evaluated for their
TCP-friendliness~\cite{friendlysurvey}, or in other words, for how
well they share a contended network resource with traditional TCP
flows as another TCP would.  By default, RemyCCs do not
play well with TCP on the same bottleneck---they are too deferential
and end up receiving less than their fair share.\footnote{RemyCCs
  generally try to keep standing queues small, both because we provide
  Remy with an objective that penalizes excess delay, but also because
  the shorter the queueing delay, the quicker the convergence time and
  the greater the throughput available to new flows. A long-running
  traditional TCP flow will work to fill queues until they overflow.}

This can be corrected, by telling Remy explicitly to assume a nonzero
risk that its algorithm may be contending with a traditional TCP
flow. In that case, the resulting RemyCCs do manage to claim their
fair share of the link (Figure~\ref{fig:tcpheterog}). But this
TCP-friendliness comes at a cost. By making the RemyCC more aggressive
in order to match up with TCP, the protocol ends up creating more
of a standing queue when it shares the network only
with other RemyCCs of the same type. From Figure~\ref{fig:tcphomog}:

\begin{center}
\includegraphics[width=0.75 \textwidth]{homo-3.pdf}
\end{center}

In other words, TCP-friendliness is a beneficial property \emph{when
  traditional TCP flows may be present on the same bottleneck}. But it
imposes a measurable cost in ``clean-slate'' designs. On internal networks
under the control of one entity, this may be worth taking into account.

\subsection{Assumptions about link rates and multiplexing are important}

We used Remy as a formalized model of the protocol-design process, to
ask rigorously: how faithfully do protocol designers really need to
understand the networks they design for?

We found that when modeling the network and link layers, it was
important for the designer's assumptions about \emph{some} network
parameters---the bottleneck link rate and the maximum degree of
multiplexing at the bottleneck---to include the true values of those
parameters.

In one experiment, we asked Remy to generate four different
congestion-control protocols, given different amounts of uncertainty
in the prior assumptions about the link rate: two-fold uncertainty,
ten-fold, hundred-fold, or thousand-fold. We then ran each protocol over a
broad range of actual link rates.

We found that performance fell dramatically when the link rate drifted
outside an algorithm's ``assumed'' range. It was possible for a
Remy-generated algorithm to outperform traditional algorithms over the
full thousand-fold range---but \textbf{only} when this possibility had
been included in the algorithm's prior assumptions. Algorithms with
more limited imagination fared poorly. From Figure~\ref{fig:breadth}:

\begin{center}
\includegraphics[width=0.75 \columnwidth]{oprange-manual.pdf}
\end{center}

The results suggest that human-generated protocols are also sensitive
to prior assumptions---the performance of Cubic-over-sfqCoDel trailed
off at higher link rates, suggesting this scheme may be making an implicit
assumption about the bandwidth-delay product that no longer holds at
such rates. However, it would be challenging to state explicitly the
assumptions of a mechanism like Cubic-over-sfqCoDel.

Among Remy-designed schemes, we also found considerable sensitivity to
prior assumptions about the maximum degree of multiplexing at a
bottleneck link. By contrast, we found less sensitivity to prior
assumptions about the expected distribution of round-trip times or the
topology of the network (Section~\ref{ss:topological}).

@@

\section{Congestion control for all}

Since the 1980s, congestion control has been a well-studied topic in
the networking community. Its importance to the Internet may still be
growing. Google is putting congestion control in its Chrome Web
browser, as part of the QUIC project to send Web traffic over UDP and
to allocate limited network resources in a way that optimizes
application-specific objectives (e.g., the Web ``speed index'').

Netflix, said to account for the majority of Internet traffic in the
United States, is working on improving its congestion control in the
context of rate selection for progressive-download video, after
research findings that its existing algorithm could produce poor results.~\cite{huang2012confused}

Meanwhile, ``big data'' batch processing in specialized datacenters
continues to explode. Optimizing the communications of computers
connected by a datacenter switching fabric---again, in order to
achieve application-specific objectives---has drawn considerable
research and commercial effort~\cite{dctcp, pfabric}.

I believe that application needs and network behaviors will continue
to evolve as the Internet keeps growing and maturing. Congestion
control, and the transport layer generally, could adapt gracefully to
this innovation if they were merely a \emph{function} of a designer's
network assumptions and application-specific goals. Sprout and Remy
represent the first steps of showing what that would look
like.

### 2014-05-21 learn_operational.tex
Clean up figure reference and add reference to contest results
Finish intro?

\begin{comment}
\begin{figure*}[b!]
%\centering
%\begin{subfigure}[b]{\columnwidth}
%\includegraphics[width=\columnwidth]{compatible-5sec.pdf}
%\caption{TCP-Aware and TCP-Naive RemyCCs, 5 second ON/OFF}
%\end{subfigure}
%
%\begin{subfigure}[b]{\columnwidth}
\caption{RemyCCs designed with and without TCP-awareness,
  competing against themselves or against TCP. Shown here, two
  endpoints contending for a 10~Mbps link with 100~ms RTT, 250 kB of
  buffer capacity (200~ms maximum queueing delay), and
  almost-continuous offered load. The fair share of throughput is
  5~Mbps per endpoint. Ellipses show 1-$\sigma$ range of results.}
%\end{subfigure}
\label{fig:tcpaware}
\begin{center}
\includegraphics[width=\textwidth]{figures/manual/compatibility-drawn.pdf}
\end{center}
\end{figure*}
\end{comment}

@@

This question has considerable practical relevance; in practice, the
developer of a new network protocol will rarely be able to arrange a
``flag day'' when all endpoints switch to the new
protocol.\footnote{There has not been a ``flag day'' on the Internet
  since the switch to IP in 1983.}

On the broad Internet today, cross-traffic will typically be the
product of traditional loss-triggered TCP congestion-control
protocols, such as NewReno or Cubic. This status quo presents a
serious problem for new protocols that seek to perform differently or
that try to avoid building up standing queues inside the network.

@@

\begin{figure}
\caption{Performance of two flows of the same congestion-control
  scheme contending for a 10~Mbps link, for three different
  congestion-control schemes. The ``TCP-naive'' RemyCC is designed
  under the assumption that its competition, when present, is governed
  by the same RemyCC---an assumption that holds here. The
  ``TCP-aware'' RemyCC is designed under the assumption that its cross
  traffic has a 50\% chance of being governed by a TCP AIMD-like
  scheme. The results quantify the degree to which designing a
  protocol to play well with TCP can exact a cost when TCP is absent.}
\label{fig:tcphomog}
\begin{center}
\includegraphics[width=\textwidth]{homo-3.pdf}
\end{center}
\end{figure}

\begin{figure}
\caption{When cross traffic is governed by TCP NewReno, the TCP-aware
  RemyCC achieves a more equitable division of the link---and better
  delay for both flows---than the TCP-naive RemyCC. The results
  demonstrate the benefit of TCP-awareness when TCP is present.}
\label{fig:tcpheterog}
\begin{center}
\includegraphics[width=\textwidth]{hetero-3.pdf}
\end{center}
\end{figure}

@@

The results suggest that new delay-minded protocols \emph{can} avoid
being squeezed out by traditional loss-triggered TCP, but building in
this behavior comes at a cost to performance in the absence of TCP
cross-traffic.

>>>>

\begin{comment}
\begin{figure}[b!]
%\centering
%\begin{subfigure}[b]{\columnwidth}
%\includegraphics[width=\columnwidth]{compatible-5sec.pdf}
%\caption{TCP-Aware and TCP-Naive RemyCCs, 5 second ON/OFF}
%\end{subfigure}
%
%\begin{subfigure}[b]{\columnwidth}
\caption{RemyCCs designed with and without TCP-awareness,
  competing against themselves or against TCP. Shown here, two
  endpoints contending for a 10~Mbps link with 100~ms RTT, 250 kB of
  buffer capacity (200~ms maximum queueing delay), and
  almost-continuous offered load. The fair share of throughput is
  5~Mbps per endpoint. Ellipses show 1-$\sigma$ range of results.}
%\end{subfigure}
\begin{center}
\includegraphics[width=\textwidth]{figures/manual/compatibility-drawn.pdf}
\end{center}
\label{fig:tcpaware}
\end{figure}
\end{comment}

@@

This question has considerable practical relevance: outside of an internal network controlled by a small number of parties, the
developer of a new network protocol will rarely be able to arrange a
``flag day'' when all endpoints switch to the new
protocol.\footnote{There has not been a ``flag day'' on the Internet
  since the switch to IP in 1983.}

On the broad Internet today, cross-traffic will typically be the
product of traditional loss-triggered TCP congestion-control
protocols, such as NewReno or Cubic. This status quo presents a
problem for new protocols that seek to perform differently or
that try to avoid building up standing queues inside the network.

@@

\begin{figure}
\caption{Performance of two flows of the same congestion-control
  scheme contending for a 10~Mbps link, for three different
  congestion-control schemes. The ``TCP-naive'' RemyCC is designed
  under the assumption that its competition, when present, is governed
  by the same RemyCC---an assumption that holds here. The
  ``TCP-aware'' RemyCC is designed under the assumption that its cross
  traffic has a 50\% chance of being governed by a TCP AIMD-like
  scheme. The results quantify the degree to which designing a
  protocol to play well with TCP can exact a cost when TCP is absent.}
\label{fig:tcphomog}
\begin{center}
\includegraphics[width=\textwidth]{homo-3.pdf}
\end{center}
\label{fig:tcpawarehomo}
\end{figure}

\begin{figure}
\caption{When cross traffic is governed by TCP NewReno, the TCP-aware
  RemyCC achieves a more equitable division of the link---and better
  delay for both flows---than the TCP-naive RemyCC. The results
  demonstrate the benefit of TCP-awareness when TCP is present.}
\label{fig:tcpheterog}
\begin{center}
\includegraphics[width=\textwidth]{hetero-3.pdf}
\label{fig:tcpawarehetero}
\end{center}
\end{figure}

@@

The results indicate that new delay-minded protocols \emph{can} avoid
being squeezed out by traditional loss-triggered TCP, but building in
this behavior comes at a cost to performance in the absence of TCP
cross-traffic.
